---
title: "Prediction for 12 session"
author: "Prakash Paudyal"
date: "6/14/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r,echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
library(caret)
set.seed(702)
known.dat<-read.csv("known.csv",header = TRUE)[,-c(1,3,4)]
trainIndx<-createDataPartition(known.dat$subject,p=0.7,list=FALSE)
train.dat<-known.dat[trainIndx,]
test.dat<-known.dat[-trainIndx,]
```

# Random Forest

```{r,echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
library(knitr)
library(randomForest)
set.seed(702)
rf.model<-randomForest(subject~.,data = train.dat, ntree=500, importance=TRUE)
rf.pred.train<-predict(rf.model,newdata=train.dat)
rf.train.acc<-mean(rf.pred.train==train.dat$subject)
#rf.train.acc
#test error
rf.pred.test<-predict(rf.model,newdata=test.dat)
rf.test.acc<-mean(rf.pred.test==test.dat$subject)
#rf.test.acc
#importance(rf.model)
#varImpPlot(rf.model)
```


#12 session

```{r,echo=FALSE,warning=FALSE,message=FALSE}
sessionn1 <- read.csv("q.session.1.csv",header = TRUE)
sessionn2 <- read.csv('q.session.2.csv',header = TRUE)
sessionn3 <- read.csv('q.session.3.csv',header = TRUE)
sessionn4<- read.csv('q.session.4.csv',header = TRUE)
sessionn5 <- read.csv('q.session.5.csv',header = TRUE)
sessionn6 <- read.csv('q.session.6.csv',header = TRUE)
sessionn7 <- read.csv('q.session.7.csv',header = TRUE)
sessionn8 <- read.csv('q.session.8.csv',header = TRUE)
sessionn9 <- read.csv('q.session.9.csv',header = TRUE)
sessionn10 <- read.csv('q.session.10.csv',header = TRUE)
sessionn11 <- read.csv('q.session.11.csv',header = TRUE)
sessionn12 <- read.csv('q.session.12.csv',header = TRUE)

```

# Prediction

# session1

subject s002 is predicted 3 times out of 9 time .The accuracy is 33.3%
Table1 for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses1<-table(predict(rf.model,newdata=sessionn1))
which.max(rf.pred.ses1)
#s002
```

```{r,echo=FALSE,warning=FALSE}
ses1.pred = table(predict(rf.model, sessionn1))
tab1 = data.frame(cbind(ses1.pred))
colnames(tab1)= "Predicted"
tab1
accuracy1 = 3/sum(diag(ses1.pred))
accuracy1
```


#session2

In session 2 subject s008 was predicted .The accuracy is 50%
Table for number of times each subject predicted is given below:


```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses2<-table(predict(rf.model,newdata=sessionn2))
which.max(rf.pred.ses2)
```

```{r,echo=FALSE,warning=FALSE}
ses2.pred = table(predict(rf.model, sessionn2))
tab2 = data.frame(cbind(ses2.pred))
colnames(tab2)= "Predicted"
tab2
accuracy1 = 3/sum(diag(ses2.pred))
accuracy1
```


#session3

In session 3 subject s020 was predicted .The accuracy is 70%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses3<-table(predict(rf.model,newdata=sessionn3))
which.max(rf.pred.ses3)
```

```{r,echo=FALSE,warning=FALSE}
ses3.pred = table(predict(rf.model, sessionn3))
tab3 = data.frame(cbind(ses3.pred))
colnames(tab3)= "Predicted"
tab3
accuracy3 = 7/sum(diag(ses3.pred))
accuracy3
```


# session 4

In session 4 subject s024 was predicted .The accuracy is 93.75%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses4<-table(predict(rf.model,newdata=sessionn4))
which.max(rf.pred.ses4)
```

```{r,echo=FALSE,warning=FALSE}
ses4.pred = table(predict(rf.model, sessionn4))
tab4= data.frame(cbind(ses4.pred))
colnames(tab4)= "Predicted"
tab4
accuracy4 = 15/sum(diag(ses4.pred))
accuracy4
```

# session 5

In session 5 subject s034 was predicted .The accuracy is 20%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses5<-table(predict(rf.model,newdata=sessionn5))
which.max(rf.pred.ses5)
```

```{r,echo=FALSE,warning=FALSE}
ses5.pred = table(predict(rf.model, sessionn5))
tab5= data.frame(cbind(ses5.pred))
colnames(tab5)= "Predicted"
tab5
accuracy5 = 2/sum(diag(ses5.pred))
accuracy5
```


# session 6

In session 6 subject s036 was predicted .The accuracy is 100%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses6<-table(predict(rf.model,newdata=sessionn6))
which.max(rf.pred.ses6)
```

```{r,echo=FALSE,warning=FALSE}
ses6.pred = table(predict(rf.model, sessionn6))
tab6= data.frame(cbind(ses6.pred))
colnames(tab6)= "Predicted"
tab6
accuracy6 = 8/sum(diag(ses6.pred))
accuracy6
```

# session 7

In session 7 subject s039 was predicted .The accuracy is 63%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses7<-table(predict(rf.model,newdata=sessionn7))
which.max(rf.pred.ses7)
```

```{r,echo=FALSE,warning=FALSE}
ses7.pred = table(predict(rf.model, sessionn7))
tab7= data.frame(cbind(ses7.pred))
colnames(tab7)= "Predicted"
tab7
accuracy7 = 12/sum(diag(ses7.pred))
accuracy7
```

# session 8

In session 8 subject s042 was predicted .The accuracy is 80%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses8<-table(predict(rf.model,newdata=sessionn8))
which.max(rf.pred.ses8)
```

```{r,echo=FALSE,warning=FALSE}
ses8.pred = table(predict(rf.model, sessionn8))
tab8= data.frame(cbind(ses8.pred))
colnames(tab8)= "Predicted"
tab8
accuracy8= 4/sum(diag(ses8.pred))
accuracy8
```

# session 9

In session 9 subject s026 was predicted .The accuracy is 33.33%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses9<-table(predict(rf.model,newdata=sessionn9))
which.max(rf.pred.ses9)
```

```{r,echo=FALSE,warning=FALSE}
ses9.pred = table(predict(rf.model, sessionn9))
tab9= data.frame(cbind(ses9.pred))
colnames(tab9)= "Predicted"
tab9
accuracy9= 2/sum(diag(ses9.pred))
accuracy9
```

# session 10

In session 10 subject s022 was predicted .The accuracy is 71.42857%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses10<-table(predict(rf.model,newdata=sessionn10))
which.max(rf.pred.ses10)
```

```{r,echo=FALSE,warning=FALSE}
ses10.pred = table(predict(rf.model, sessionn10))
tab10= data.frame(cbind(ses10.pred))
colnames(tab10)= "Predicted"
tab10
accuracy10= 5/sum(diag(ses10.pred))
accuracy10
```

# session 11

In session 11 subject s044 was predicted .The accuracy is 78%
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses11<-table(predict(rf.model,newdata=sessionn11))
which.max(rf.pred.ses11)
```

```{r,echo=FALSE,warning=FALSE}
ses11.pred = table(predict(rf.model, sessionn11))
tab11= data.frame(cbind(ses11.pred))
colnames(tab11)= "Predicted"
tab11
accuracy11= 39/sum(diag(ses11.pred))
accuracy11
```

# Session 12

In session 12 subject s048 was predicted .The accuracy is 25 %
Table for number of times each subject predicted is given below:

```{r,echo=FALSE,warning=FALSE,message=FALSE}

rf.pred.ses12<-table(predict(rf.model,newdata=sessionn12))
which.max(rf.pred.ses12)
```

```{r,echo=FALSE,warning=FALSE}
ses12.pred = table(predict(rf.model, sessionn12))
tab12= data.frame(cbind(ses12.pred))
colnames(tab12)= "Predicted"
tab12
accuracy12= 7/sum(diag(ses12.pred))
accuracy12
```


# Prediction Table

```{r}
library(knitr)
sessionIndex<-c(1:12)
predicted<-c("s002","s008","s020","s024" ,"s034","s036","s039","s042","s031","s022","s044","s048")
Accuracy<-c(33.33333,50,70,93.75,20,100,15.78947,80,33.33333,71.42857,78,25)
finltable<-data.frame(sessionIndex,predicted,Accuracy)
#kable(finltable)
Table1<-kable(finltable,caption = "Table for prediction Acuracy")
Table1
```